Why do languages tolerate heterography? An experimental investigation into the emergence of informative orthography

### 2024-01-15 manuscript/manuscript.tex
Fix typo

It is particularly surprising that participants chose not to innovate novel systems in the communicative condition, despite their being a clear financial incentive to do so. Why might this be the case? Firstly, although innovation was not expressly forbidden in the instructions, it was perhaps suggested by the overall structure of the experiment that participants were expected to use the orthography learned during training. Participants who adopted this interpretation may have feared diverging from what they had learned, either because they themselves wanted to avoid breaking the rules or because they did not know whether their partner would feel comfortable breaking the rules. Secondly, even if participants were willing to diverge, they were presented with an alignment problem: How should they decide what system of color marking to use when there is no channel through which they can agree a system with their partners? They can only use trial-and-error, and even then, the process is complicated by participants simultaneously adapting to each other's systems. These are dynamics that also exist in the real world, especially in writing where the reader and writer are separated in time and space. Additionally, deviation from spelling norms is often considered socially unacceptable or uneducated, even when novel spellings are advantageous (e.g., the non-standard spelling <thru> has struggled for decades to make any headway against <through>). Even under more progressive attitudes to change and innovation, the problem of achieving community buy-in remains, especially when the change is not imposed by top-down authority: If someone decided to start using the spelling <banque> to differentiate the financial institution from river banks, would anyone take that spelling seriously or even understand the intention? Differentiation is hard in the written modality.

>>>>

It is particularly surprising that participants chose not to innovate novel systems in the communicative condition, despite there being a clear financial incentive to do so. Why might this be the case? Firstly, although innovation was not expressly forbidden in the instructions, it was perhaps suggested by the overall structure of the experiment that participants were expected to use the orthography learned during training. Participants who adopted this interpretation may have feared diverging from what they had learned, either because they themselves wanted to avoid breaking the rules or because they did not know whether their partner would feel comfortable breaking the rules. Secondly, even if participants were willing to diverge, they were presented with an alignment problem: How should they decide what system of color marking to use when there is no channel through which they can agree a system with their partners? They can only use trial-and-error, and even then, the process is complicated by participants simultaneously adapting to each other's systems. These are dynamics that also exist in the real world, especially in writing where the reader and writer are separated in time and space. Additionally, deviation from spelling norms is often considered socially unacceptable or uneducated, even when novel spellings are advantageous (e.g., the non-standard spelling <thru> has struggled for decades to make any headway against <through>). Even under more progressive attitudes to change and innovation, the problem of achieving community buy-in remains, especially when the change is not imposed by top-down authority: If someone decided to start using the spelling <banque> to differentiate the financial institution from river banks, would anyone take that spelling seriously or even understand the intention? Differentiation is hard in the written modality.
